---
title: PTAPP 2017, IAAO SURVEY
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: readable
---
# Introduction

Hi, welcome to PTAPP! Currently would need to differentiate the two Hawaii answers as separate jurisdictions, due to jurisdiction_name values.

```{r setup, include=FALSE} 
# Remove warnings from outputting to consumer
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo = F, include = F}
# necessary packages
library(DBI)
library(odbc)
library(dplyr)
library(ggplot2)
library(tidyr)
library(config)
library(Rcpp)
library(rmarkdown)
library(knitr)
library(kableExtra)
library(rgdal)
library(leaflet)
library(plotly)
library(leaflet.extras)
library(DT)
library(stringr)
library(geojsonio)
#_________ Ingest data 
# # connect to database
source('functions.R') # to grab the histogram generator function
# having trouble with connections on my current local PC; tabling this and opting for text file in this demo
importedPtapp <- read.table('ptapp_data.txt', header = T, sep = "\t")

#_________ Clean data
ptapp <- importedPtapp %>% rename(Response = response) %>% rename(Question = sub_question) %>% rename(Jurisdiction = jurisdiction_name)
is.na(ptapp) <- ptapp == 'NULL'

#________ Choropleth setup from https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
# download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="world_shape_file.zip")
# You now have it in your current working directory, have a look!

#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
# Read this shape file with the rgdal library. 
world_spdf <- readOGR( 
  dsn= paste0(getwd(),"/world_shape_file") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# Fake participation indicator for choropleth proof of concept:
world_spdf@data <- world_spdf@data %>% dplyr::mutate(participation = dplyr::case_when(
  NAME == 'Canada' ~ 1,
  NAME == 'United States' ~ 1,
  NAME == 'China' ~ 1,
  NAME == 'United Kingdom' ~ 1,
  NAME == 'South Korea' ~ 1,
  NAME == 'Netherlands' ~ 1,
  NAME == 'South Africa' ~ 1,
  T ~ NA_real_
))

world_spdf@data <- world_spdf@data %>% dplyr::mutate(participationNames = dplyr::case_when(
 participation == 1 ~ NAME,
 T ~ ''
))

myParticipationPalette <- colorNumeric("darkred", domain = world_spdf@data$participation, na.color = 'transparent')

# State geojson download for sample US map (probably not realistic - would need an international map with canadian provinces, us states, etc.)
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")

```

# International Representativeness
Great turnout internationally this year. 

```{r, echo = F}
map <- leaflet::leaflet(data = world_spdf) %>% addTiles() %>% addPolygons(fillColor = ~myParticipationPalette(participation), stroke = F, fillOpacity = 0.8, label = world_spdf@data$participationNames) %>% leaflet.extras::addFullscreenControl() %>% leaflet::setMaxBounds(lng1 = -180, lng2 = 180, lat1 = -90, lat2 = 90) %>% leaflet::setView( lat=0, lng=0 , zoom=1.25)
map
```

# Survey Limitations
Just as an example to tie data into the choropleth, since the international map did not coincide with the national US state jurisdictions pulled as example data for this exercise.

This map colors the jurisdiction polygons by the number of total questions & nested sub-questions responded to.

```{r, echo = F}
ptapp_state_join <- ptapp %>% group_by(Jurisdiction) %>% summarize(num_responses = sum(responded_to)) %>% rename(name = Jurisdiction)

statesForMap <- merge(states,ptapp_state_join, by.x = 'name')
bins <- c(0, 25, 35, 40, 45, 55, 60, 65, Inf)
pal <- colorBin("YlOrRd", domain = statesForMap$num_responses, bins = bins)
labels <- paste0(
  statesForMap$name, "<br/> Num. Unique Question Responses: ", statesForMap$num_responses
) %>% lapply(htmltools::HTML)

m <- leaflet(statesForMap) %>%
  setView(-96, 37.8, 3) %>%
  addTiles() %>% addPolygons(fillColor = ~pal(num_responses), weight = 2, opacity = 1, color = 'white', dashArray = 3, fillOpacity = 0.7,highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
    label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto")) %>% addLegend(pal = pal, values = ~density, opacity = 0.7, title = 'Num. Responses',
  position = "bottomright")

m
```

Below is a table of the number of missing responses per question. If one question had 10 nested questions, and one jurisdiction did not answer, that equates to 10 missing responses. For this exercise, responded_to = 0 was randomly assigned and not translated from IAAO's true PTAPP 2017 data.

```{r, echo = F}
no_responses <- ptapp %>% filter(responded_to == 0) %>% group_by(question_num) %>% rename(`Question Number` = question_num) %>% summarize(`Number of Missing Responses` = n())

# Table
knitr::kable(x = no_responses, format = 'html', row.names = F, caption = 'Count of Unanswered Questions, Multiplied by Num. Sub Questions') %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '100%')
```

```{r, echo = F, include = F}
## Data Map
#All responses embedded into map's polygons is a great idea but would best be managed through ESRI or open-source technology dedicated to such representations. This was running into issues with SPDF type objects and it did not feel productive or clean to widen question-subquestion responses into new columns.


# ptapp_state_join2 <- ptapp %>%  select(c(Jurisdiction, question_num, question_desc, Question, Response)) %>% rename(name = Jurisdiction)
# 
# statesForMap <- merge(states,ptapp_state_join2, by.x = 'name', duplicateGeoms = T)
# 
# 
# labs <- lapply(seq(nrow(statesForMap)), function(i) {
#   paste0( '<p>', statesForMap[i, "name"], '<p></p>', 
#           statesForMap[i, "question_num"], ', ', 
#           statesForMap[i, "question_desc"],'</p><p>', 
#           statesForMap[i, "Question"],
#           statesForMap[i, "Response"], '<br/>',
#           '</p>' ) 
# })
# 
# 
# 
# 
# m <- leaflet(statesForMap) %>%
#   setView(-96, 37.8, 3) %>%
#   addTiles() %>% addPolygons(fillColor = 'darkblue', weight = 2, opacity = 1, color = 'white', dashArray = 3, fillOpacity = 0.7,highlightOptions = highlightOptions(
#     weight = 5,
#     color = "#666",
#     dashArray = "",
#     fillOpacity = 0.7,
#     bringToFront = TRUE),
#     label = labels,
#   labelOptions = labelOptions(
#     style = list("font-weight" = "normal", padding = "3px 8px"),
#     textsize = "15px",
#     direction = "auto"))
# 
# m
```


# Question 1
```{r, echo = F}
#______ Question 1
# Grab only question 1 rows
q1 <- ptapp %>% dplyr::filter(question_num == 1)
# Obtain columns for graphics
if(!is.na(q1$Question[1])) {
  q1_data <- q1 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q1_data <- q1 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q1_header <- q1 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q1_data, format = 'html', row.names = T, caption = q1_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
``` 

## Comments
More responses than received in 2012's edition of PTAPP.

# Question 2
```{r, echo = F}
#______ Question 2
# Grab only question 2 rows
q2 <- ptapp %>% dplyr::filter(question_num == 2)
# Obtain columns for graphics
if(!is.na(q2$Question[1])) {
  q2_data <- q2 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q2_data <- q2 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q2_header <- q2 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q2_data, format = 'html', row.names = T, caption = q2_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

## Comments
Placeholder text.

# Question 3
```{r, echo = F}
#______ Question 3
# Grab only question 3 rows
q3 <- ptapp %>% dplyr::filter(question_num == 3)
# Obtain columns for graphics
if(!is.na(q3$Question[1])) {
  q3_data <- q3 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q3_data <- q3 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q3_header <- q3 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q3_data, format = 'html', row.names = T, caption = q3_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  %>% scroll_box(width = '100%', height = '350px')
```
## Comments
Placeholder text.

# Question 4

Complete response data listed in the table below.

```{r, echo = F}
#______ Question 4
# Grab only question 4 rows
q4 <- ptapp %>% dplyr::filter(question_num == 4)
# Obtain columns for graphics
if(!is.na(q4$Question[1])) {
  q4_data <- q4 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q4_data <- q4 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q4_header <- q4 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q4_data, format = 'html', row.names = T, caption = q4_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

```{r, echo = F}
q4_hist_data <- q4_data %>% filter(Question == 'Size Status')
q4_hist <- ggplot2::ggplot(q4_hist_data, aes(Response)) + ggplot2::geom_histogram(stat = "count", position = "stack", binwidth = 1) + theme_minimal() + ggplot2::ggtitle(q4_hist_data$Question)
q4_hist
```

Demonstrations of ggplot2 vs plotly for same data:

```{r, echo = F}
q4_freq_data <- q4_data %>% filter(Question == 'Number of staff members in the office')
q4_freq <- ggplot2::ggplot(q4_freq_data, aes(Jurisdiction,as.numeric(Response))) + ggplot2::geom_col() + theme_minimal() + ggplot2::ggtitle(q4_freq_data$Question) + ylab('Number of Staff') + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) 
q4_freq

q4_freq_plotly = plotly::plot_ly(
  x = q4_freq_data$Jurisdiction,
  y = as.numeric(q4_freq_data$Response),
  type = 'bar'
) %>% 
  layout(title = q4_freq_data$Question, xaxis = list(title = 'Jurisdiction'), yaxis = list(title = 'Staff Count'))

q4_freq_plotly
```

## Comments
Placeholder text. Notice the plotly vs ggplot2 graph differences (plotly is interactive). There is concern bundling many plotlys as their interactivity may slow down the document. Effects here are TBD.


# Question 5

Raw data:

```{r, echo = F}
#______ Question 5
# Grab only question 5 rows
q5 <- ptapp %>% dplyr::filter(question_num == 5)
# Obtain columns for graphics
if(!is.na(q5$Question[1])) {
  q5_data <- q5 %>% select(c("Jurisdiction", "Question", "Response", "question_desc"))

} else {
 q5_data <- q5 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q5_header <- q5 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q5_data, format = 'html', row.names = T, caption = q5_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

The following graphs display the response distribution of the `r length(unique(q5_data$Question))` Sub Questions for Question `r q5$question_num[1]`. 

```{r, echo = F}

for (entry in 1:length(unique(q5_data$Question))) {
  q5_data_hist <- q5_data %>% filter(Question == unique(Question)[entry])
  hist_func(q5_data_hist)
}

```

## Comments
Wow, great job people!


# Question 6
Complete response data listed in the table below.

```{r, echo = F}
#______ Question 6
# Grab only question 6 rows
q6 <- ptapp %>% dplyr::filter(question_num == 6)
# Obtain columns for graphics
if(!is.na(q6$Question[1])) {
  q6_data <- q6 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q6_data <- q6 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q6_header <- q6 %>% dplyr::select(question_desc) %>% unique()
# Numeric Response, so convert it:
q6_data$Response <- as.numeric(q6_data$Response)
# Table
knitr::kable(x = q6_data, format = 'html', row.names = T, caption = q6_header, format.args = list( big.mark = ",")) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

## Comments
Lots of parcels. 

# Question 7
Complete response data listed in the table below. This uses the datatable/DT framework rather than kable framework for tables. Note the differences.

```{r, echo = F}
#______ Question 7
# Grab only question 7 rows
q7 <- ptapp %>% dplyr::filter(question_num == 7)
# Obtain columns for graphics
if(!is.na(q7$Question[1])) {
  q7_data <- q7 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q7_data <- q7 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q7_header <- q7 %>% dplyr::select(question_desc) %>% unique()
# Numeric Response, so convert it:
q7_data$Response <- as.numeric(q7_data$Response) %>% format(big.mark=',')
# Table
DT::datatable(q7_data, 
              options = list(pageLength = 20, scrollX = T, scrollY = T, order = list(list(1, 'desc'))), 
              class = 'cell-border stripe',
              caption = htmltools::tags$caption(
              style = 'caption-side: top; text-align: center;',
              q7_header),
              rownames = F)
```

## Comments
Poor response rate for this question. Does this imply a lack of personal property in the equalization process for these jurisdictions?

# Question 8
Complete response data listed in the table below.

```{r, echo = F}
#______ Question 8
# Grab only question 8 rows
q8 <- ptapp %>% dplyr::filter(question_num == 8)
# Obtain columns for graphics
if(!is.na(q8$Question[1])) {
  q8_data <- q8 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q8_data <- q8 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q8_header <- q8 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q8_data, format = 'html', caption = q8_header, row.names = T) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

The following graphs display the response distribution of the `r length(unique(q8_data$Question))` Sub Questions for Question `r q8$question_num[1]`. 

```{r, echo = F}
for (entry in 1:length(unique(q8_data$Question))) {
  q8_data_hist <- q8_data %>% filter(Question == unique(Question)[entry])
  hist_func(q8_data_hist)
}
```

## Comments
Lorem ipsum

# Question 9
Complete response data listed in the table below.

```{r, echo = F}
#______ Question 9
# Grab only question 9 rows
q9 <- ptapp %>% dplyr::filter(question_num == 9)
# Obtain columns for graphics
if(!is.na(q9$Question[1])) {
  q9_data <- q9 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q9_data <- q9 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q9_header <- q9 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q9_data, format = 'html', caption = q9_header, row.names = T) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

The following graphs display the response distribution of the `r length(unique(q9_data$Question))` Sub Questions for Question `r q9$question_num[1]`. 

```{r, echo = F}
for (entry in 1:length(unique(q9_data$Question))) {
  q9_data_hist <- q9_data %>% filter(Question == unique(Question)[entry])
  hist_func(q9_data_hist)
}
```

## Comments
Lorem ipsum

# Question 10
Complete response data listed in the table below.

```{r, echo = F}
#______ Question 10
# Grab only question 10 rows
q10 <- ptapp %>% dplyr::filter(question_num == 10)
# Obtain columns for graphics
if(!is.na(q10$Question[1])) {
  q10_data <- q10 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q10_data <- q10 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q10_header <- q10 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q10_data, format = 'html', caption = q10_header, row.names = T) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

The following graphs display the response distribution of the `r length(unique(q10_data$Question))` Sub Questions for Question `r q10$question_num[1]`. 

```{r, echo = F}
for (entry in 1:length(unique(q10_data$Question))) {
  q10_data_hist <- q10_data %>% filter(Question == unique(Question)[entry])
  hist_func(q10_data_hist)
}
```

## Comments
Lorem ipsum