---
title: PTAPP 2017, IAAO SURVEY
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: readable
---
# Introduction

Hi, welcome to PTAPP! Currently would need to differentiate the two Hawaii answers as separate jurisdictions, due to jurisdiction_name values.

```{r setup, include=FALSE} 
# Remove warnings from outputting to consumer
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo = F, include = F}
# necessary packages
library(DBI)
library(odbc)
library(dplyr)
library(ggplot2)
library(tidyr)
library(config)
library(Rcpp)
library(rmarkdown)
library(knitr)
library(kableExtra)
library(rgdal)
library(leaflet)
library(plotly)
library(leaflet.extras)
#_________ Ingest data 
# # connect to database
# source('functions.R')
# having trouble with connections on my current local PC; tabling this and opting for text file in this demo
importedPtapp <- read.table('ptapp_data.txt', header = T, sep = "\t")

#_________ Clean data
ptapp <- importedPtapp %>% rename(Response = response) %>% rename(Question = sub_question) %>% rename(Jurisdiction = jurisdiction_name)
is.na(ptapp) <- ptapp == 'NULL'

#________ Choropleth setup from https://r-graph-gallery.com/183-choropleth-map-with-leaflet.html
# Download the shapefile. (note that I store it in a folder called DATA. You have to change that if needed.)
# download.file("http://thematicmapping.org/downloads/TM_WORLD_BORDERS_SIMPL-0.3.zip" , destfile="world_shape_file.zip")
# You now have it in your current working directory, have a look!

#  -- > You now have 4 files. One of these files is a .shp file! (TM_WORLD_BORDERS_SIMPL-0.3.shp)
# Read this shape file with the rgdal library. 
world_spdf <- readOGR( 
  dsn= paste0(getwd(),"/world_shape_file") , 
  layer="TM_WORLD_BORDERS_SIMPL-0.3",
  verbose=FALSE
)

# Fake participation indicator for choropleth proof of concept:
world_spdf@data <- world_spdf@data %>% dplyr::mutate(participation = dplyr::case_when(
  NAME == 'Canada' ~ 1,
  NAME == 'United States' ~ 1,
  NAME == 'China' ~ 1,
  NAME == 'United Kingdom' ~ 1,
  NAME == 'South Korea' ~ 1,
  NAME == 'Netherlands' ~ 1,
  NAME == 'South Africa' ~ 1,
  T ~ NA_real_
))

world_spdf@data <- world_spdf@data %>% dplyr::mutate(participationNames = dplyr::case_when(
 participation == 1 ~ NAME,
 T ~ ''
))

myParticipationPalette <- colorNumeric("darkred", domain = world_spdf@data$participation, na.color = 'transparent')
```

# International Representativeness
Great turnout internationally this year. 

```{r, echo = F}
map <- leaflet::leaflet(data = world_spdf) %>% addTiles() %>% addPolygons(fillColor = ~myParticipationPalette(participation), stroke = F, fillOpacity = 0.8, label = world_spdf@data$participationNames) %>% leaflet.extras::addFullscreenControl()
map
```

# Question 1
```{r, echo = F}
#______ Question 1
# Grab only question 1 rows
q1 <- ptapp %>% dplyr::filter(question_num == 1)
# Obtain columns for graphics
if(!is.na(q1$Question[1])) {
  q1_data <- q1 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q1_data <- q1 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q1_header <- q1 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q1_data, format = 'html', row.names = T, caption = q1_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
``` 

## Comments
More responses than received in 2012's edition of PTAPP.

# Question 2
```{r, echo = F}
#______ Question 2
# Grab only question 2 rows
q2 <- ptapp %>% dplyr::filter(question_num == 2)
# Obtain columns for graphics
if(!is.na(q2$Question[1])) {
  q2_data <- q2 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q2_data <- q2 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q2_header <- q2 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q2_data, format = 'html', row.names = T, caption = q2_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

## Comments
Placeholder text.

# Question 3
```{r, echo = F}
#______ Question 3
# Grab only question 3 rows
q3 <- ptapp %>% dplyr::filter(question_num == 3)
# Obtain columns for graphics
if(!is.na(q3$Question[1])) {
  q3_data <- q3 %>% select(c("Jurisdiction", "Question", "Response"))
} else {
 q3_data <- q3 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q3_header <- q3 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q3_data, format = 'html', row.names = T, caption = q3_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))  %>% scroll_box(width = '100%', height = '350px')
```
## Comments
Placeholder text.

# Question 4

Complete response data listed in the table below.

```{r, echo = F}
#______ Question 4
# Grab only question 4 rows
q4 <- ptapp %>% dplyr::filter(question_num == 4)
# Obtain columns for graphics
if(!is.na(q4$Question[1])) {
  q4_data <- q4 %>% select(c("Jurisdiction", "Question", "Response"))

} else {
 q4_data <- q4 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q4_header <- q4 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q4_data, format = 'html', row.names = T, caption = q4_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

```{r, echo = F}
q4_hist_data <- q4_data %>% filter(Question == 'Size Status')
q4_hist <- ggplot2::ggplot(q4_hist_data, aes(Response)) + ggplot2::geom_histogram(stat = "count", position = "stack", binwidth = 1) + theme_minimal() + ggplot2::ggtitle(q4_hist_data$Question)
q4_hist
```

Demonstrations of ggplot2 vs plotly for same data:

```{r, echo = F}
q4_freq_data <- q4_data %>% filter(Question == 'Number of staff members in the office')
q4_freq <- ggplot2::ggplot(q4_freq_data, aes(Jurisdiction,as.numeric(Response))) + ggplot2::geom_col() + theme_minimal() + ggplot2::ggtitle(q4_freq_data$Question) + ylab('Number of Staff') + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) 
q4_freq

q4_freq_plotly = plotly::plot_ly(
  x = q4_freq_data$Jurisdiction,
  y = as.numeric(q4_freq_data$Response),
  type = 'bar'
) %>% 
  layout(title = q4_freq_data$Question, xaxis = list(title = 'Jurisdiction'), yaxis = list(title = 'Staff Count'))

q4_freq_plotly
```

## Comments
Placeholder text. Notice the plotly vs ggplot2 graph differences (plotly is interactive). There is concern bundling many plotlys as their interactivity may slow down the document. Effects here are TBD.


# Question 5

Raw data:

```{r, echo = F}
#______ Question 5
# Grab only question 5 rows
q5 <- ptapp %>% dplyr::filter(question_num == 5)
# Obtain columns for graphics
if(!is.na(q5$Question[1])) {
  q5_data <- q5 %>% select(c("Jurisdiction", "Question", "Response", "question_desc"))

} else {
 q5_data <- q5 %>% select(c("Jurisdiction", "Response"))
}
# Obtain Question Description programmatically for reproducibility + header
q5_header <- q5 %>% dplyr::select(question_desc) %>% unique()
# Table
knitr::kable(x = q5_data, format = 'html', row.names = T, caption = q5_header) %>% kableExtra::kable_paper() %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = '100%', height = '350px')
```

Distribution of Answers to each Sub Question:

```{r, echo = F}
q5_hist_func <- function(x) { 
  q5_hist <- ggplot2::ggplot(filter(q5_data, q5_data$Question == x), aes(Response)) + ggplot2::geom_bar() + theme_minimal() + ggplot2::labs(title = q5_data$question_desc, subtitle = x)
  return(q5_hist)
  }
  
lapply(unique(q5_data$Question), q5_hist_func)

#   
# for (entry in 1:length(unique(q5_data$Question))) {
#   q5_data <- q5_data %>% filter(q5_data$Question == unique(q5_data$Question)[entry])
#   q5_hist_func(q5_data$Question)
# }

```
